---
title: "Untitled"
author: "Mai D."
date: "11/26/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview


The Student Performance dataset is downloaded from Kaggle Student Performance in Exams <https://www.kaggle.com/spscientist/students-performance-in-exams> consists of 1000 observations and 8 columns - variables, the scores are marked to the students in 3 subjects: Math, Reading, and Writing:

Gender
Ethnicity
Parental level of education
Lunch
Test preparation course
Math score
Reading score
Writing score

The inspiration is to understand the influence of the student's background and determine the features which play an important role in affecting academic performance. 


## Dataset

```{r read}
# Read the dataset
data <- read.csv("/Users/maido/Google Drive/Asgm/Statistical Programming/Assignment/StudentsPerformance.csv", header = TRUE)
```

# Library

```{r lib}
library(dplyr)
library(ggplot2)
library(gridExtra)
```

Combining the three scores into average score as a meansure of student perfomance

```{r comb}
# Average score in 3 subjects
avg.score <- rowSums(data[ , 6:8])/3
data <- cbind(data, avg.score)

# Preview of the dataset
head(data)
summary(data)
```


## Visualisation

Firstly plotting the class labels and see the differences in the plot. Thus, we can determine which ones are the most beneficial values and which ones are not.

# Comparison of Gender attributes to the Marks

```{r boxplot, echo=FALSE}
par(mfrow=c(2,2))
boxplot(data$math.score~data$gender, col=rainbow(7), main="Math", xlab="Gender", ylab="Math Score")
boxplot(data$reading.score~data$gender, col=rainbow(2), main="Reading", xlab="Gender", ylab="Reading Score")
boxplot(data$writing.score~data$gender, col=rainbow(3), main="Writing", xlab="Gender", ylab="Writing Score")
boxplot(data$avg.score~data$gender, col=rainbow(4), main="Average", xlab="Gender", ylab="Average Score")
```

comment

However, we need to use Hypothesis tests are designed to detect whether an "effect" is systematic or is the result of random variation. 

```{r hypotest1}
# Comparing the means of Math by Gender – the unpaired samples t-test
var.test(data$math.score~data$gender)
```

The p-value is greater than 0.05, so do not reject the null hypothesis that the variances are equal. 


```{r hypo2}
# Perform an unpaired samples t-test using
t.test(data$math.score~data$gender, var.equal=TRUE)
```

We cannot reject the null hypothesis H0. There is some difference in the math score of male and female


# Comparison of Race Ethnicity attributes to the Marks

```{r race1, echo=FALSE}
# Mean of Group
math.score <- data %>% group_by(race.ethnicity) %>% summarise(math.avg = mean(math.score), .groups = 'drop')
reading.score <- data %>% group_by(race.ethnicity) %>% summarise(reading.avg = mean(reading.score), .groups = 'drop')
writing.score <- data %>% group_by(race.ethnicity) %>% summarise(writing.avg = mean(writing.score), .groups = 'drop')
avg.score <- data %>% group_by(race.ethnicity) %>% summarise(avg = mean(avg.score), .groups = 'drop')

# Plot to compare 
par(mfrow=c(2,2))
barplot(math.score$math.avg~math.score$race.ethnicity,
        main="Average of Math Score",
        xlab="Race", ylab="Average Math Scores")
barplot(reading.score$reading.avg~reading.score$race.ethnicity,
        main="Average of Reading Score",
        xlab="Race", ylab="Average Reading Scores")
barplot(writing.score$writing.avg~writing.score$race.ethnicity,
        main="Average of Writing Score",
        xlab="Race", ylab="Average Writing Scores")
barplot(avg.score$avg~avg.score$race.ethnicity,
        main="Average Score",
        xlab="Race", ylab="Average Scores")
```

comment


# Comparison of Test Preparation attributes to the Marks

```{r test1, echo=FALSE}
library(gridExtra)
g1 <- ggplot(subset(data), aes(x=math.score, colour=test.preparation.course)) +
  geom_density() +
  ggtitle("Math Score")
g2 <- ggplot(subset(data), aes(x=reading.score, colour=test.preparation.course)) +
  geom_density() +
  ggtitle("Reading Score")
g3 <- ggplot(subset(data), aes(x=writing.score, colour=test.preparation.course)) +
  geom_density() +
  ggtitle("Writing Score")
g4 <- ggplot(subset(data), aes(x=avg.score, colour=test.preparation.course)) +
  geom_density() +
  ggtitle("Average Score")
grid.arrange(g1, g2, g3, g4, nrow = 2, ncol  = 2)
```

comment

# Hypothesis Testing

```{r pretest}
# Comparing the means of Reading by Pre-Test Course
var.test(data$reading.score~data$test.preparation.course)
```

The p-value is greater than 0.05, so do not reject the null hypothesis that the variances are equal. 


```{r prest, echo=FALSE}
# Perform an unpaired samples t-test using
t.test(data$reading.score~data$test.preparation.course, var.equal=TRUE)
```

We cannot reject the null hypothesis H0. There is some difference in the reading score of preparing course


# Comparison of Parent's Education attributes to the Average Marks

```{r parent, echo=FALSE}
ggplot(data=data, mapping=aes(x=parental.level.of.education, y = avg.score, 
                              col=parental.level.of.education))+
  theme_bw() +
  geom_boxplot()+
  scale_y_continuous(limits=c(0,110),breaks = seq(0,110,10))+
  labs(title="Parent's Education effects the Average Marks", 
       x="Parent's Education Status", y="Average Score")
```



```{r ANOVA}
# ANOVA test to compare parents education and average marks
ANOVA<- aov(data$avg.score ~ data$parental.level.of.education)
summary(ANOVA)
```


## Decsriptive Statistics

# Summary on Pre-Test Course determining the average score

```{r press, echo=FALSE}
data %>% group_by(test.preparation.course) %>% summarise(
  Min = min(avg.score,na.rm = TRUE),
  Q1 = quantile(avg.score,probs= .25,na.rm = TRUE),
  Median = median(avg.score, na.rm = TRUE),
  Q3 = quantile(avg.score,probs = .75,na.rm = TRUE),
  Max = max(avg.score,na.rm = TRUE),
  Mean = mean(avg.score, na.rm = TRUE),
  SD = sd(avg.score, na.rm = TRUE),
  n = n(),
  Missing = sum(is.na(avg.score)), .groups = 'drop')
```


# Summary on gender determining the average score

```{r prf, echo=FALSE}
data %>% group_by(gender) %>% summarise(
  Min = min(avg.score,na.rm = TRUE),
  Q1 = quantile(avg.score,probs= .25,na.rm = TRUE),
  Median = median(avg.score, na.rm = TRUE),
  Q3 = quantile(avg.score,probs = .75,na.rm = TRUE),
  Max = max(avg.score,na.rm = TRUE),
  Mean = mean(avg.score, na.rm = TRUE),
  SD = sd(avg.score, na.rm = TRUE),
  n = n(),
  Missing = sum(is.na(avg.score)), .groups = 'drop')
```


# Summary on parents determining the average score

```{r pressdgs, echo=FALSE}
data %>% group_by(parental.level.of.education) %>% summarise(
  Min = min(avg.score,na.rm = TRUE),
  Q1 = quantile(avg.score,probs= .25,na.rm = TRUE),
  Median = median(avg.score, na.rm = TRUE),
  Q3 = quantile(avg.score,probs = .75,na.rm = TRUE),
  Max = max(avg.score,na.rm = TRUE),
  Mean = mean(avg.score, na.rm = TRUE),
  SD = sd(avg.score, na.rm = TRUE),
  n = n(),
  Missing = sum(is.na(avg.score)), .groups = 'drop')
```


# Summary on Race Ethnicity determining the average score

```{r presds, echo=FALSE}
data %>% group_by(gender) %>% summarise(
  Min = min(avg.score,na.rm = TRUE),
  Q1 = quantile(avg.score,probs= .25,na.rm = TRUE),
  Median = median(avg.score, na.rm = TRUE),
  Q3 = quantile(avg.score,probs = .75,na.rm = TRUE),
  Max = max(avg.score,na.rm = TRUE),
  Mean = mean(avg.score, na.rm = TRUE),
  SD = sd(avg.score, na.rm = TRUE),
  n = n(),
  Missing = sum(is.na(avg.score)), .groups = 'drop')
```



## Other questions

# Students are good at math and bad at writing? 

```{r presh, echo=FALSE}
data <- data %>% mutate(diff = data$math.score - data$writing.score)
barplot(c(sum(data$diff < 0), sum(data$diff > 0)), 
        main="Does student good at math bad at writing?", 
        xlab="No - Yes", ylab="count")
```


# T - test

```{r prev, echo=FALSE}
diff <- data$math.score - data$writing.score
mean(diff)
t.test(diff, mu=0)
```

Comment: p-value is less than 0.05 so we conclude that the difference is 0. Thus we can conclude there is no enough evidence to prove that the students who do good is math is bad at writing.

